
---
title: "Cheatsheet for tidyr function"

output: github_document
---
## Main Functions

 tidyr Function | Special Powers
-------------------|----------------------------
  `gather`         | long format\*
  `spread`         | wide format\*
  `separate()`     | Splitting a single variable into two\*
  `unite()`        | Merging two variables into one\*
```{r}
library(tidyr)
```

### `gather` function

![](http://b243.photo.store.qq.com/psb?/752a85a3-4499-4558-bfba-dc2edb3ff4fe/.B7chYYghbbwZcF0s1itOjzghh4M7XLe61LDo2oGg9g!/c/dPMAAAAAAAAA&bo=8AN0AvADdAIDACU!&rf=mood_app)

**Description** the gather() function will take multiple columns into key-value pairs, duplicating all other columns as needed.

Function:       gather(data, key, value, ..., na.rm = FALSE, convert = FALSE)

Arguments:
        data:           data frame
        key:            column name representing new variable
        value:          column name representing variable values
        ...:            names of columns to gather (or not gather)
        na.rm:          option to remove observations with missing values (represented by NAs)
        convert:        if TRUE will automatically convert values to logical, integer, numeric, complex or 
                        factor as appropriate
                        
#### Example

```{r}
year<-c(2010:2017)

Qtr.1<-c(10:17)
Qtr.2<-c(20:27)
Qtr.3<-c(30:37)
data<-data.frame(year,Qtr.1,Qtr.2,Qtr.3)
knitr::kable(data)
```

```{r}

new <-data %>%
  gather(Quarter, value, Qtr.1:Qtr.3)

knitr::kable(new)
```

### `seperate()` function

**Description**  Many times a single column variable will capture multiple variables, so `seperate` function  will split a single variable into two.

Function:       separate(data, col, into, sep = " ", remove = TRUE, convert = FALSE)
Same as:        data %>% separate(col, into, sep = " ", remove = TRUE, convert = FALSE)

Arguments:
        data:           data frame
        col:            column name representing current variable
        into:           names of variables representing new variables
        sep:            how to separate current variable (char, num, or symbol)
        remove:         if TRUE, remove input column from output data frame
        convert:        if TRUE will automatically convert values to logical, integer, numeric, complex or 
                        factor as appropriate
 
         
#### Example
                       
```{r}
new2 <- new %>% separate(Quarter, c("Time_Interval", "Interval_ID"))

knitr::kable(new2)
```

### `seperate()` function

**Description** There may be a time in which we would like to combine the values of two variables, so `unite` function will merge two variables into one.

Function:       unite(data, col, ..., sep = " ", remove = TRUE)
Same as:        data %>% unite(col, ..., sep = " ", remove = TRUE)

Arguments:
        data:           data frame
        col:            column name of new "merged" column
        ...:            names of columns to merge
        sep:            separator to use between merged values
        remove:         if TRUE, remove input column from output data frame

        
#### Example
        
```{r}
new3 <- new2 %>% unite(Quarter, Time_Interval, Interval_ID, sep = ".")

knitr::kable(new3)

```

### `spread( )` function:

**Description**  There are times when we are required to turn long formatted data into wide formatted data. The `spread()` function spreads a key-value pair across multiple columns.

Function:       spread(data, key, value, fill = NA, convert = FALSE)
Same as:        data %>% spread(key, value, fill = NA, convert = FALSE)

Arguments:
        data:           data frame
        key:            column values to convert to multiple columns
        value:          single column values to convert to multiple columns' values 
        fill:           If there isn't a value for every combination of the other variables and the key 
                        column, this value will be substituted
        convert:        if TRUE will automatically convert values to logical, integer, numeric, complex or 
                        factor as appropriate

#### Example

```{r}
new4<- spread(new, Quarter,value )


knitr::kable(new4)
```

## Related link

+ [tidyr](https://rpubs.com/bradleyboehmke/data_wrangling)

+ [tidyr2](https://blog.rstudio.com/2014/07/22/introducing-tidyr/)

+ [tidyr3](https://www.rdocumentation.org/packages/tidyr/versions/0.7.1/topics/gather)