---
title: "Expolre gapminder"

output: github_document
---
## Bring rectangular data in

### First time to install the data `gapminder` and `dplyr`
```{r}
##install.packages("gapminder")

##install.packages("tidyverse")
```

### Load the data
```{r}
library(gapminder)
library(tidyverse)
```

## Smell test the data
**Q1: Is it a data.frame, a matrix, a vector, a list?**


```{r}
typeof(gapminder)
```
From the output, `gapminder` data is a list. 
**Q2: What’s its class?**
```{r}
class(gapminder)
```

From the R output above, it shows that gapminder should be a data.frame, but also a tibble.

**Q3: How many variables/columns?**
```{r}
ncol(gapminder)

length(gapminder)
```
There are two ways to show how many variables using `ncol` and `length` functions.
In addition, we can print `gapminder` dirctly by using `head` or `tail` function.
```{r}
head(gapminder)
tail(gapminder)
```

**Q4: How many rows/observations?**
```{r}
nrow(gapminder)

```
There are 1704 observations.
**Q5: Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?**
```{r}
dim(gapminder)
```
One way to shows how many rows and variables is using `dim()`.
From the output, `gapminder` has 1704 rows and 6 columns
If we need to deal with the vector, at that time we can not use `dim()` function, because if different columns are linear dependent, the number of columns not equal to the dimension of columns.
More methods about `gapminder` size.
```{r}
summary(gapminder)
str(gapminder)
```
From `summary` function, we can find the mean, median, minimum, maximum about each variable. And using `str`, we could find name, class, type of data.  
**Q6: What data type is each variable?**

```{r}
class(gapminder$country)
class(gapminder$continent)
class(gapminder$year)
class(gapminder$lifeExp)
class(gapminder$pop)
class(gapminder$gdpPercap)
```
The second way
```{r}
typeof(gapminder$country) 
typeof(gapminder$continent) 
typeof(gapminder$year)
typeof(gapminder$lifeExp)
typeof(gapminder$pop)
typeof(gapminder$gdpPercap)
```
The third way
```{r}
str(gapminder)
```

## Explore individual variables
There are 6 variables for gapminder, 3 quantitative variables: `year`, `lifeExp`, `gdpPercap` and `pop`; 2 categorical variable: `conti nent` and `country`.

**Q1: What are possible values (or range, whichever is appropriate) of each variable?**
##### Categorical variables #####
The categorial variable `continent`.
```{r}
##The first method
summary(gapminder$continent)
##The secound method
gapminder %>% 
  distinct(continent)
##the third method
  unique(gapminder$continent)
```
There are 5 value for `continent` : Africa, Americas, Asia, Europe and Oceania.
the Categorial variable `country`.

```{r}
##The first method
summary(gapminder$country)
##The secound method
gapminder %>% 
  distinct(country)
##the third method
  unique(gapminder$country)
```
### Quantitative variable
#### Life Expectancy variable
```{r}
head(gapminder$lifeExp)
summary(gapminder$lifeExp)
hist(gapminder$lifeExp)

```

From the R out put the min for Life Expectancy is 23.6
the largest value for Life Expectancy is 82.6
The data is right skewed
70 to 75 Life Expctancy have the highest frequency

#### popualtion variable
```{r}
head(gapminder$pop)
summary(gapminder$pop)
hist(gapminder$pop, freq = F)

```

## Explore various plot types

### A scatterplot of two quantitative variables.
```{r}
new<- gapminder %>% 
  filter(continent==c("Europe", "Asia", "Americas")) %>% 
  select(continent,year,lifeExp,gdpPercap) %>% 
  rename(life_exp = lifeExp,
         gdp_percap = gdpPercap)

```

```{r}
p<- ggplot(new, aes(x = gdp_percap, y = life_exp))
p + geom_point(aes(color=continent,size=year), alpha = 0.5)+scale_x_log10()
```


