---
title: "Expolre gapminder"

output: github_document
---
## Bring rectangular data in

### First time to install the data `gapminder` and `dplyr`
```{r}
##install.packages("gapminder")

##install.packages("tidyverse")
```

### Load the data
```{r}
library(gapminder)
library(tidyverse)
```

## Smell test the data

**Q1: Is it a data.frame, a matrix, a vector, a list?**


```{r}
typeof(gapminder)
```
From the output, `gapminder` data is a list. 
**Q2: What’s its class?**
```{r}
class(gapminder)
```

From the R output above, it shows that gapminder should be a data.frame, but also a tibble.

**Q3: How many variables/columns?**
```{r}
ncol(gapminder)

length(gapminder)
```
There are two ways to show how many variables using `ncol` and `length` functions.
In addition, we can print `gapminder` dirctly by using `head` or `tail` function.
```{r}
head(gapminder)
tail(gapminder)
```

**Q4: How many rows/observations?**
```{r}
nrow(gapminder)

```
There are 1704 observations.

**Q5: Can you get these facts about “extent” or “size” in more than one way? Can you imagine different functions being useful in different contexts?**
```{r}
dim(gapminder)
```
One way to shows how many rows and variables is using `dim()`.
From the output, `gapminder` has 1704 rows and 6 columns
If we need to deal with the vector, at that time we can not use `dim()` function, because if different columns are linear dependent, the number of columns not equal to the dimension of columns.
More methods about `gapminder` size.
```{r}
summary(gapminder)
str(gapminder)
```
From `summary` function, we can find the mean, median, minimum, maximum about each variable. And using `str`, we could find name, class, type of data.  

**Q6: What data type is each variable?**

```{r}
class(gapminder$country)
class(gapminder$continent)
class(gapminder$year)
class(gapminder$lifeExp)
class(gapminder$pop)
class(gapminder$gdpPercap)
```
The second way
```{r}
typeof(gapminder$country) 
typeof(gapminder$continent) 
typeof(gapminder$year)
typeof(gapminder$lifeExp)
typeof(gapminder$pop)
typeof(gapminder$gdpPercap)
```
The third way
```{r}
str(gapminder)
```

## Explore individual variables
There are 6 variables for gapminder, 3 quantitative variables: `year`, `lifeExp`, `gdpPercap` and `pop`; 2 categorical variable: `conti nent` and `country`.

**Q1: What are possible values (or range, whichever is appropriate) of each variable?**

##### Categorical variables 
The categorial variable `continent`.
```{r}
##The first method
summary(gapminder$continent)
##The secound method
gapminder %>% 
  distinct(continent)
##the third method
  unique(gapminder$continent)
```
There are 5 value for `continent` : Africa, Americas, Asia, Europe and Oceania.
the Categorial variable `country`.

```{r}
##The first method
summary(gapminder$country)
##The secound method
gapminder %>% 
  distinct(country)
##the third method
  unique(gapminder$country)
```
##### Quantitative variable
The Quantitative variable `lifeExp`.
```{r}
summary(gapminder$lifeExp)
```

```{r}
## Another way using %>%
gapminder %>% 
  summarize(mean_lifeExp=mean(lifeExp), min_lifeExp=min(lifeExp), max_lifeExp=max(lifeExp), median_lifeExp=median(lifeExp))
```
The range for life expectancy between 23.6 and 83.6, the mean is 59.47, median is 60.71.


The Quantitative variable `pop`.
```{r}
summary(gapminder$pop)
```
```{r}
## Another way using %>%
gapminder %>% 
  summarize(mean_lifeExp=mean(pop), min_lifeExp=min(pop), max_lifeExp=max(pop), median_lifeExp=median(pop))
```

The range for population from 60011 and 1318683096, the mean is 29601212, median is 7023596.

** Q2: What values are typical? What’s the spread? What’s the distribution? Etc., tailored to the variable at hand.**
We usually use mean to represent typical value, and we can estimate spread using variance or IQR. Moreover, we can plot the graph to find the distribution.

##### Categorical variables 
The categorial variable `continent`.
```{r}
summary(gapminder$continent)
table(gapminder$continent)
barplot(table(gapminder$continent))
```
From above, we can find how many data in different continent. Asia should be a typical value, because Asia has the largest data in `gapminder`. From the plot, it shows how spread of the `continent`.
the Categorial variable `country`.
```{r}
table(gapminder$country)
barplot(table(gapminder$country))
```
From above, we can find 12 data for each country. And from the barplot, the `country` variable should have uniform distribution. 

##### Quantitative variable
The Quantitative variable `lifeExp`.
```{r}
summary(gapminder$lifeExp)
##another way
gapminder %>% 
  summarize(mean_lifeExp=mean(lifeExp), sd_lifeExp=sd(lifeExp), min_lifeExp=min(lifeExp), max_lifeExp=max(lifeExp), median_lifeExp=median(lifeExp))
hist(gapminder$lifeExp)

```
The mean for lifeExp is 59.47. Consider about spread, if we use standard diviation: sd =12.9. From the plot above, it can clearly see that the distribution skewed to the right. Most of the data fall in the interval(65,80).

The Quantitative variable `pop`.
```{r}
summary(gapminder$pop)
##another way
gapminder %>% 
  summarize(mean_lifeExp=mean(pop), sd_lifeExp=sd(pop), min_lifeExp=min(pop), max_lifeExp=max(pop), median_lifeExp=median(pop))
hist(gapminder$pop)

```
The mean for lifeExp is 29601212. Consider about spread, if we use standard diviation: sd =106157897. From the plot above, it can clearly see that the distribution skewed to the left. Most of the data fall in the interval(0, 10^8).


## Explore various plot types

+ **A scatterplot of two quantitative variables.**

#### Scatterplot 1: scatterplot of `lifeExp` and `gdpPercap`, only consider the continent: Asia, Europe and Americas
Using `filter`, `select`, and `rename` dplyr.
```{r}
new<- gapminder %>% 
  filter(continent==c("Europe", "Asia", "Americas")) %>% 
  select(continent,year,lifeExp,gdpPercap) %>% 
  rename(life_exp = lifeExp,
         gdp_percap = gdpPercap)

```
Ggplot based on new dataset, add the color for continent, size for year. this plot changes transparency to 0.5, and make a log transformation.
```{r}
p1<- ggplot(new, aes(x = gdp_percap, y = life_exp))
p1 + geom_point(aes(color=continent,size=year), alpha = 0.5)+scale_x_log10()
```

#### Scatterplot 2: Scatterplot of 'year' and 'pop', only consider three countries: United states, Canada, China 

```{r}
new2<- gapminder %>% 
  filter(country==c("France", "Canada", "China")) %>% 
  select(country,continent,year,pop)  
```
Adding shape for country, color for continent, also using `geom_line`
```{r}
p2 <- ggplot(new2, aes(year, pop))
p2 + geom_point(aes(shape=country, color= continent))+
geom_line(aes(group = country))
```
#### Scatterplot 3: Scatterplot of 'year' and 'lifeExp'
Using `geom_jitter` and `geom_smooth` 
```{r}
ggplot(gapminder, aes(x = year, y = lifeExp,
                      color = continent)) +
  geom_jitter(alpha = 0.5, size = 2)+
    geom_smooth(lwd = 3, se = FALSE)
```
Seperate different continent by using `facet_wrap()`
```{r}
ggplot(gapminder, aes(x = year, y = lifeExp,
                      color = continent)) +
  facet_wrap(~ continent) +
geom_jitter(alpha = 0.5, size = 2)+
 geom_smooth(lwd = 2, se = F)
```

+ **A plot of one quantitative variable. Maybe a histogram or densityplot or frequency polygon.**

